### 2019-06-21 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done proofreading quickly Ch5

As we saw in Chapters~\ref{chapter:1} and \ref{chapter:4},
a crucial step for the development of Cognitive Radio is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
%
In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
Such networks can be modelled using a decentralized multi-players multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
%
The decentralized hypothesis means that the central base station does not given any direct order to the devices, they are all in charge of deciding the channel they each want to use at each time step.
Moreover, the devices are independent and do not directly communicate with each other either.
We show that dynamic end-devices are able to use limited feedback sent by the gateway to learn to find orthogonal affectations of the group of end-devices to the best radio channels, automatically and without explicit communication with each other.
In other words, the end-devices are able to learn to cooperate efficiently, in a completely decentralized and autonomous manner.

@@

\textbf{Outline.}
%
The rest of this chapter is organized as follows.
% First we motivate the model of multi-players multi-armed bandits in Section~\ref{sec:5:motivation}
We introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}.
We present a useful regret decomposition in Section~\ref{sec:5:lowerbound}, and illustrate it in the Appendix~\ref{sec:5:appendix}.
The \Selfish, \RandTopM{} and \MCTopM{} algorithms are introduced in Section~\ref{sec:5:algorithms},
for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
The result of our experimental study reported in Section~\ref{sec:5:experiments}.
Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
% and finally Section~\ref{sec:5:conclusion} concludes.

\textbf{Publication.}
%
This chapter is mainly based on our article \cite{Besson2018ALT}.

@@

Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
we use the Kullback-Leibler divergence to express our lower bound.
We remind that $\kl(x,y) \eqdef x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes the assumption of uniform efficiency given in Definition~\ref{def:2:uniformlyEfficientAlgorithm},
a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
% Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
% This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
%
% \vspace*{-20pt}  % XXX
\begin{definition}\label{def:5:DecentralizedUniformEfficiency}
\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
  An algorithm $\cA$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$,
  % and for all $\alpha \in (0,1)$,
  % \vspace*{-5pt}  % XXX
  \begin{align}
    \forall a\in(0,1) & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
    \text{and} \forall a\in(0,1) & \;
    \forall j \in [M], k \in \Mbest, \;\;
    % a_{j,k}
    \frac{T}{M}
    - \E_{\mu}[N_k^j(T)] \mathop{=}\limits_{T \to +\infty} o(T^{\alpha}).
    \label{eq:5:SUE}
  \end{align}
\end{leftbar}  % XXX leftbar defnbar, comment if needed
\end{definition}

Having a small regret on every problem instance, \ie, being uniformly efficient,
is a natural assumption,
that rules out algorithms tuned to perform well on specific instances only (\eg, fixed-armed algorithms).
%
From this assumption $\left(R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)\right)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
${T}- \E_{\mu}[N_k(T)] {=} o(T^{\alpha})$, and so
\begin{equation}\label{eq:5:intermediate}
  % \vspace*{-5pt}  % XXX
  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[N_k^j(T)]\right) = o(T^{\alpha}).
  % \vspace*{-5pt}  % XXX
\end{equation}

@@

% As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
The notion of fairness defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is given and known to the algorithm.

@@

The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} come from Appendix~E of \cite{KaufmannAbbas19}.
Similarly to what we did in Section~\ref{sec:2:notations},
consider the observations $O_t^j$ that she gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
defined as $O_t^j \eqdef \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$ (\eg, for an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index, the decision is an $\argmax$ that is usually a uniform random selection among the arms with maximum index).

@@

This result can also be used to recover Proposition~1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
the \emph{expected number of sub-optimal selections} -- $\E_{\mu}[N_k(T)]$ for $k \in \Mworst$ --
and the \emph{expected number of colliding players on the optimal arms} -- $\E_{\mu}[\cC_k(T)]$ for $k \in \Mbest$. Note that, in term (c) the number of colliding players on the sub-optimal arm $k$ may be upper bounded as $\E_{\mu}[\cC_k(T)] \leq M \E_{\mu}[N_k(T)]$.
%

In the next section, we present an algorithm that has a logarithmic regret,
while ensuring that the number of sub-optimal selections is matching the lower bound of Theorem~\ref{thm:5:BetterLowerBound}.

@@

\textcolor{red}{\textbf{Warning:}}
We want to strengthen the fact that in all the proposed algorithms,
the players do \emph{not} know their numbers $j\in[M]$,
and they do not need to know it to achieve low regret.
%
We discuss the fact that this would be an unrealistic hypothesis in Section~\ref{par:5:knowingYourIDinMPBanditsGame} below.

@@

% \textbf{Index policies.}
% %
The approache we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
$\UCB_1$ \citep{Auer02}
and \klUCB{} \citep{KLUCBJournal}.
For player $j\in[M]$
%
denote $S_k^j(t) \eqdef \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/N_k^j(t)$ (if $N_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  U_k^j(t) \eqdef \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2N_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: N_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

\textbf{Two ideas for orthogonalization.}
\label{par:5:twoIdeasOrthogonalization}
%
On the one hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.
%
On the other hand, the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}. %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst. Extending our algorithms for the case of unknown number of players $M$ is an interesting future work.}.
%
More formally, letting $\pi(k,\mathbf{g})$ be the index of the $k$-th largest entry in a vector $\mathbf{g}$,
% (for $k\in[K]$),
in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in[M]$
and selects at time $t$,
% \begin{equation}
$A^j(t) \eqdef \pi\left(R^j(t), [U^j_\ell(t)]_{\ell=1,\dots,K}\right)$.
% \end{equation}
If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.

% Now, our algorithms !!
\textbf{Our two proposals.}
We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
\begin{equation}
  \TopM(t) \eqdef \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell=1,\dots,K}\right), k=1,\dots,M\right\}.
\end{equation}

\textbf{The \RandTopM{} algorithm.}
%
We precisely state our first proposal below in Algorithm~\ref{algo:5:RandTopM}.

@@

\RandTopM{} is essentially a refinement over \RhoRand, to not use the indirection of ranks, and a simpler version of \MCTopM.
The difference with \MCTopM{} is that the later introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
%
In \RandTopM{}, player $j$ is always considered ``not fixed'',
and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.

\textbf{The \MCTopM{} algorithm.}
%
Our second proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
it is a slightly more complex extension of the \RandTopM{} algorithm.
From there on, we focus on \MCTopM{} as it is easier to analyze and performs better.
%
Both algorithms ensure that player $j$ always
selects at time $t+1$ an arm from $\TopM(t)$.
When a collision occurs for a player implementing the \RandTopM{} algorithm, that player randomly switches arm within $\TopM$, while \MCTopM{} uses a more sophisticated mechanism, that is reminiscent of ``Musical Chair'' (MC) and inspired by the work of \cite{Rosenski16}: players tend to fix themselves on arms (``chairs'') and ignore future collision when this happens.

@@

  As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t-1)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
  Formally, $s^j(0)$ is false, and $s^j(t)$ is defined inductively from $s^j(t-1)$ as
  % following the algorithm of \MCTopM:
  \begin{equation}
      s^j(t) =
      \left( s^j(t-1) \cup \left( \overline{s^j(t-1)} \cap \overline{C^j(t-1)} \right) \right)
      \cap \left( A^j(t) \in \TopM(t-1) \right).
  \end{equation}

  For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words. At step $t$, if player $j$ is not fixed ($\overline{s^j(t-1)}$), she can have three behaviors when executing \MCTopM.
  She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
  or she stays in state $\overline{s^j(t)}$,
  with two cases.
  Either she sampled $A^j(t)$ uniformly
  from $\TopM(t) \cap \{ m : U_m^j(t-1) \leq U_k^j(t-1) \}$
  with transition $(3)$,
  in case of collision and if $A^j(t-1) \in \TopM(t)$,
  or she sampled $A^j(t)$ uniformly
  from $\TopM(t)$ with transition $(2)$,
  if $A^j(t-1) \notin \TopM(t)$.
  In particular, note that if $\overline{C^j(t-1)}$, transition $(3)$ is executed and not $(2)$.
  %
  % The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
  % and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
  % and experienced a collision at time $t-1$
  Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).

@@

  To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
  we introduce three sequences of random variables,
  the starting times $(\theta_i)_{i \geq 1}$
  and the ending times $(\tau_i)_{i \geq 1}$
  (possibly larger than $T$),
  of sequences during which $I_2(s)$ is true for all $s=\theta_i,\dots,\tau_i-1$ but not before and after,
  that is
  $\forall i \in \{1,\dots,n(T)\},
  \overline{I_2^j(\theta_i - 1)}
  \cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
  \cap \overline{I_2^j(\tau_i)}
  $
  with $n(T)$ the number of such sequences,
  \ie,
  $n(T) \eqdef \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
  (or $0$ if $\theta_1$ does not exist).
  %
  If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.

  Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
  \[
    \E_{\mu}[N_2^j(t)]
    = \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
    =
    \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
    =
    \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
  \]
  Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$),
  thus we can use Wald's Lemma \citep{Wald45},
  to obtain a decomposition with two terms $(\alpha)$ and $(\beta)$,
  $\E_{\mu}[N_2^j(t)] \leq \E_{\mu}\left[ n(T) \right] \times \max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]$.
  % $\E_{\mu}[N_2^j(t)] \leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}$.

>>>>

As we saw in Chapters~\ref{chapter:1} and \ref{chapter:4},
a crucial step for the development of Cognitive Radio is to insert \emph{multiple} smart devices in the \emph{same} background traffic (at least $M \geq 2$).
%
In this Chapter~\ref{chapter:5}, we are interested in a more formal approach to the decentralized learning problem presented in the previous Chapter~\ref{chapter:4}.
Such networks can be modelled using a decentralized multi-players multi-armed bandit problem, where arms are channels and players are dynamic IoT devices.
%
The decentralized hypothesis means that the central base station (or gateway) does not given any direct order to the devices, they are all in charge of deciding the channel they each want to use at each time step.
Moreover, the devices are independent and do not directly communicate with each other either.
We show that dynamic end-devices are able to use limited feedback sent by the gateway  to learn to find orthogonal affectations of the group of end-devices to the best radio channels, automatically and without explicit communication with each other.
In other words, the end-devices are able to learn to cooperate efficiently, in a completely decentralized and autonomous manner.

@@

\paragraph{Outline.}
%
The rest of this chapter is organized as follows.
% First we motivate the model of multi-players multi-armed bandits in Section~\ref{sec:5:motivation}
We introduce the multi-players bandit model with three feedback levels in Section~\ref{sec:5:model}.
We present a useful regret decomposition in Section~\ref{sec:5:lowerbound}, and illustrate it in the Appendix~\ref{sec:5:appendix}.
The \Selfish, \RandTopM{} and \MCTopM{} algorithms are introduced in Section~\ref{sec:5:algorithms},
for which we present our theoretical analysis in Section~\ref{sec:5:upperbounds}.
The result of our experimental study reported in Section~\ref{sec:5:experiments}.
Finally, we present in Section~\ref{sec:5:literatureReviewOtherModels} a review of the recent literature which studies variants of the model presented in this Chapter.
For some extensions, we discuss how to adapt our proposals and illustrate the empirical performances of such modification of \MCTopM-\klUCB{}, leaving theoretical analyses as future works.
% and finally Section~\ref{sec:5:conclusion} concludes.

\paragraph{Publication.}
%
This chapter is mainly based on our article \cite{Besson2018ALT}.

@@

Similarly to what we present above in Section~\ref{sec:2:lowerUpperBoundsRegret},
we use the Kullback-Leibler divergence to express our lower bound.
We remind that $\kl(x,y) \eqdef x\log(x/y) + (1-x)\log((1-x)/(1-y))$ is the KL divergence between the Bernoulli distributions of means $x$ and $y$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes the assumption of uniform efficiency given in Definition~\ref{def:2:uniformlyEfficientAlgorithm},
a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
% Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
% This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.
%
% \vspace*{-20pt}  % XXX
\begin{definition}\label{def:5:DecentralizedUniformEfficiency}
\begin{leftbar}[defnbar]  % XXX leftbar defnbar, comment if needed
  An algorithm $\cA$  is \emph{\textbf{strongly uniformly efficient}} if for all $\boldsymbol{\mu} \in \cP_M$,
  % and for all $\alpha \in (0,1)$,
  % \vspace*{-5pt}  % XXX
  \begin{align}
    \forall a\in(0,1) & R_T(\boldsymbol{\mu},M,\cA) \mathop{=}\limits_{T \to +\infty} o(T^\alpha) \\
    \text{and } \; \forall a\in(0,1) & \;
    \forall j \in [M], k \in \Mbest, \;\;
    % a_{j,k}
    \frac{T}{M}
    - \E_{\mu}[N_k^j(T)] \mathop{=}\limits_{T \to +\infty} o(T^{\alpha}).
    \label{eq:5:SUE}
  \end{align}
\end{leftbar}  % XXX leftbar defnbar, comment if needed
\end{definition}

Having a small regret on every problem instance (\ie, being uniformly efficient)
is a natural assumption,
that rules out policies tuned to perform well only on specific instances (\eg, fixed-armed policies).
%
From this assumption $R_T(\boldsymbol{\mu},M,\cA)=o(T^\alpha)$ and the decomposition of Lemma~\ref{lem:5:DecompositionRegret} one can see\footnote{With some arguments used in the proof of Lemma~\ref{lem:5:1stLowerBound} to circumvent the fact that \ref{eq:5:term2} may be negative.} that for every $k \in \Mbest$,
${T}- \E_{\mu}[N_k(T)] {=} o(T^{\alpha})$, and so
\begin{equation}\label{eq:5:intermediate}
  % \vspace*{-5pt}  % XXX
  \sum_{j =1}^M\left(\frac{T}{M} - \E_{\mu}[N_k^j(T)]\right) = o(T^{\alpha}).
  % \vspace*{-5pt}  % XXX
\end{equation}

@@

% As we state it here, the constraint on fairness is new, and it was not studied as such in previous works.
\textbf{About fairness.}
%
In the proofs of the regret bounds for \rhoRand{} in \cite{Anandkumar11}, the authors briefly mention that their policy is invariant under permutation and that this yields a certain fairness guarantee, but without formalizing it more.
The notion of fairness defined above can be interpreted as a ``cooperative fairness'', opposed to the notion of ``arm fairness'' that has been studied in a few papers on single-player stochastic MAB.
For instance, \cite{Patil2019stochastic} requires that each arm must be sampled at least a given fraction of the total time steps, \ie, $\forall k, \E[N_k(T)] \geq r_k T$, where the constraint vector $[r_1,\dots,r_K]$ is known by the algorithm.

@@

\textbf{Collision information.}
%
The following notations and the hypothesis on the collision information term \eqref{eq:5:collisionInformationTerm} comes from Appendix~E of \cite{KaufmannAbbas19}.
Similarly to what we did in Section~\ref{sec:2:notations},
consider the observations $O_t^j$ that she gathered after $t$ rounds of its algorithm $\cA_j$ for a fixed player $j\in[M]$,
defined as $O_t^j \eqdef \left( U^j(1), Y_{A^j(1),1}, C^j(1), \ldots, U^j(t), Y_{A^j(t),t}, C^j(t) \right)$,
where $U^j(t)$ denotes some external source of randomness useful to select $A^j(t+1)$.
For instance, an algorithm based on ranks and \UCB{} indexes, like \RhoRand, when two arms have the same maximum index the decision is an $\argmax$, and ties are usually broken by a uniform random selection among the arms with maximum index.

@@

This result can also be used to recover Proposition~1 from \cite{Anandkumar11}, giving an upper bound on the regret that only depends on
the \emph{expected number of sub-optimal selections}, $\E_{\mu}[N_k(T)]$ for $k \in \Mworst$,
and the \emph{expected number of colliding players on the optimal arms}, $\E_{\mu}[\cC_k(T)]$ for $k \in \Mbest$. Note that, in term (c) the number of colliding players on the sub-optimal arm $k$ may be upper bounded as $\E_{\mu}[\cC_k(T)] \leq M \E_{\mu}[N_k(T)]$.
%

In the next section, we present an algorithm that has a logarithmic regret,
by controlling its sub-optimal selections,
while ensuring that the number of sub-optimal selections is matching the lower bound of Theorem~\ref{thm:5:BetterLowerBound}.

@@

\begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
  \textcolor{red}{\textbf{Warning.}}
  We want to strengthen the fact that in all the proposed algorithms,
  the players do \emph{not} know their numbers $j\in[M]$,
  and they do not need to know it to achieve low regret.
  %
  This would be an unrealistic hypothesis, as explained in Section~\ref{par:5:knowingYourIDinMPBanditsGame} below.
\end{leftbar}  % XXX leftbar warningbar, comment if needed

@@

% \textbf{Index policies.}
% %
The approaches we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to UCB algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
$\UCB_1$ \citep{Auer02}
and \klUCB{} \citep{KLUCBJournal}.
For player $j\in[M]$
%
denote $S_k^j(t) \eqdef \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/N_k^j(t)$ (if $N_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  U_k^j(t) \eqdef \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2N_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: N_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

\paragraph{Two ideas for orthogonalization.}
\label{par:5:twoIdeasOrthogonalization}
%
On the one hand, the \TDFS{} algorithm \citep{Zhao10} relies on the player agreeing in advance on the time steps at which they will target each of the $M$ best indices.
Even though some alternative without pre-agreement are proposed, they are quite complicated and we prefer to focus on other approaches.
%
On the other hand, the \rhoRand{} algorithm \citep{Anandkumar11} relies on randomly selected \emph{ranks}. %\footnote{\rhoRand{} only works if $M$ is known, and \cite{Anandkumar11} extended it to the case of an unknown $U$, with \rhoRandEst. Extending our algorithms for the case of unknown number of players $M$ is an interesting future work.}.
%
More formally, letting $\pi(k,\mathbf{U})$ be the index of the $k$-th largest entry in a vector $\mathbf{U}$,
% (for $k\in[K]$),
in \rhoRand{} each player maintains at time $t$ an internal rank $R^j(t)\in[M]$
and selects at time $t$,
% \begin{equation}
$A^j(t) \eqdef \pi\left(R^j(t), [U^j_\ell(t)]_{\ell\in[K]}\right)$.
% \end{equation}
If a collision occurs, a new rank is drawn uniformly at random, $R^j(t+1) \sim \cU([M])$.

% Now, our algorithms !!
\paragraph{Our two proposals.}
We now propose two alternatives to this strategy, that do not rely on ranks and rather randomly fix themselves on one \emph{arm} in $\TopM(t)$, that is defined as the set of arms that have the $M$ largest indices (at the current time $t$ and for player $j$),
\begin{equation}
  \TopM(t) \eqdef \left\{ \pi\left(k, \{U^j_\ell(t)\}_{\ell\in[K]}\right), k=1,\dots,M\right\}.
\end{equation}

\paragraph{The \RandTopM{} algorithm.}
%
We precisely state our first proposal below in Algorithm~\ref{algo:5:RandTopM}.
%
\RandTopM{} is essentially a refinement over \RhoRand, to not use the indirection of ranks, and a simpler version of \MCTopM.
The difference with \MCTopM{} is that the later introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
%
In \RandTopM{}, player $j$ is always considered ``not fixed'',
and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.

@@

\paragraph{The \MCTopM{} algorithm.}
%
Our second proposal \MCTopM{} is stated below as Algorithm~\ref{algo:5:MCTopM},
it is a slightly more complex extension of the \RandTopM{} algorithm.
From there on, we focus on \MCTopM{} as it is easier to analyze and performs better.
%
Both algorithms ensure that player $j$ always
selects at time $t+1$ an arm from $\TopM(t)$.
When a collision occurs for a player implementing the \RandTopM{} algorithm, that player randomly switches arm within $\TopM$, while \MCTopM{} uses a more sophisticated mechanism, that is reminiscent of ``Musical Chair'' (MC) and inspired by the work of \cite{Rosenski16}: players tend to fix themselves on arms (``chairs'') and ignore future collision when this happens.

@@

  As in Algorithm~\ref{algo:5:MCTopM}, $s^j(t-1)$ is the event that player $j$ decided to fix herself on an arm at the end of round $t-1$.
  Thus $s^j(0)$ is false, and $s^j(t)$ is defined inductively from $s^j(t-1)$ as
  % following the algorithm of \MCTopM:
  \begin{equation}
      s^j(t) =
      \left( s^j(t-1) \cup \left( \overline{s^j(t-1)} \cap \overline{C^j(t-1)} \right) \right)
      \cap \left( A^j(t) \in \TopM(t-1) \right).
  \end{equation}

  For the sake of clarity, we now explain Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} in words. At step $t$, if player $j$ is not fixed ($\overline{s^j(t-1)}$), she can have three behaviors when executing \MCTopM.
  She keeps the same arm and goes to the other state $s^j(t)$ with transition $(1)$,
  or she stays in state $\overline{s^j(t)}$
  in two cases.
  Either she sampled $A^j(t)$ uniformly
  from $\TopM(t) \cap \{ m : U_m^j(t-1) \leq U_k^j(t-1) \}$
  with transition $(3)$,
  in case of collision and if $A^j(t-1) \in \TopM(t)$,
  or she sampled $A^j(t)$ uniformly
  from $\TopM(t)$ with transition $(2)$,
  if $A^j(t-1) \notin \TopM(t)$.
  In particular, note that $(3)$ is executed and not $(2)$ if $\overline{C^j(t-1)}$.
  %
  % The case $(2)$ happens when there is no collision at time $t-1$ but a collision at time $t$,
  % and conversely case $(3)$ happens when the player was not fixed on her arm at time $t-1$ (\ie, $\overline{s^j(t-1)}$)
  % and experienced a collision at time $t-1$
  Transition $(3)$ is a uniform sampling from $\TopM(t)$ (the ``Musical Chair'' step).

@@

  To control the first term $\sum_{t=1}^T \Pr(I_2^j(t))$
  we introduce three sequences of random variables,
  the starting times $(\theta_i)_{i \geq 1}$
  and the ending times $(\tau_i)_{i \geq 1}$
  (possibly larger than $T$),
  of sequences during which $I_2(s)$ is true for all $s \in [\theta_i,\dots,\tau_i-1]$, but not before and not after,
  that is
  $\forall i \in \{1,\dots,n(T)\},
  \overline{I_2^j(\theta_i - 1)}
  \cap \bigcap_{t=\theta_i}^{\tau_i-1} I_2^j(t)
  \cap \overline{I_2^j(\tau_i)}
  $
  with $n(T)$ the number of such sequences,
  \ie,
  $n(T) \eqdef \inf \{i \geq 1 : \min(\theta_i, \tau_i) \geq T \}$
  (or $0$ if $\theta_1$ does not exist).
  %
  If $\theta_i = 1$, the first sequence does not have term $\overline{I_2^j(\theta_i - 1)}$.

  Now we can decompose the sum on $t=1,\dots,T$ with the use of consecutive sequences,
  \[
    \E_{\mu}[N_2^j(t)]
    = \E_{\mu}\left[ \sum_{t=1}^T \indic{\left(I_2^j(t)\right)} \right]
    =
    \E_{\mu}\left[ \sum_{i=1}^{n(T)} \left( \sum_{t=\theta_i}^{\tau_i - 1} 1 + \sum_{t=\tau_i}^{\theta_{i+1} - 1} 0 \right) \right]
    =
    \E_{\mu}\left[ \sum_{i=1}^{n(T)} (\tau_i - \theta_i) \right].
  \]
  Both $n(T)$ and $\tau_i - \theta_i \geq 0$ have finite averages for any $i$ (as $\tau_i - \theta_i \leq T$), and $n(T)$ is a \emph{stopping time} with respect to the past events (that is, $\cF^j_T$),
  thus we can use Wald's Lemma \citep{Wald45},
  to obtain a decomposition with two terms $(\alpha)$ and $(\beta)$,
  $\E_{\mu}[N_2^j(t)] \leq \E_{\mu}\left[ n(T) \right] \times \max\limits_{i \geq 1} \E_{\mu}\left[ \tau_i - \theta_i \right]$.
  % $\E_{\mu}[N_2^j(t)] \leq \underbrace{\E_{\mu}\left[ n(T) \right]}_{(\alpha)} \times \underbrace{\max_{i\in\N} \E_{\mu}\left[ \tau_i - \theta_i \right]}_{(\beta)}$.
